Report on the Fourth Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE4)

### 2017-01-03 WSSSPE4_report/wg_soft_eng.tex
minor changes

%Introduction to group here, including the overall objective of work in this area.
This working group concerned itself with identifying processes that are not adequately
covered by general software engineering. Verification and testing is
the first subtopic that we address.  
% there is a growing awareness about the lack of trust
% and reproducibility in many computational results. 
%\katznote{this paragraph repeats some of the previous one - they probably should be merged}
Computational science and engineering applications have many moving
parts that need to interoperate with one another. The accuracy and
reliability of results produced by the scientific software depends not
only on the individual components behaving correctly, but also on the
validity of their interactions. 
% Therefore, a rigorous verification
% process and a robust testing regime are critical requirements for
% scientific software.  There have been instances where inadequate
% verification has resulted in publication of wrong results and later
% retraction of publications.
As scientific understanding grows
the corresponding computational software models are refined, leading
to more complex codes. Increasing complexity makes them more prone to
defects, not only in individual code units, but also in interaction
among units. Therefore, a strong verification process combined with a
rigorous testing regime plays a critical role in the prevention of
generating incorrect scientific results. However, most science teams
struggle to find a good solution for themselves. Causes range from
lack of exposure to the practices, to distrust of adopting practices
because they do not meet the needs of the teams developing such
software.  The current focus of our effort is on testing because it is the first step
towards building a software processes that can lead to provenance and
reproducibility, the hallmarks of quality science.

>>>>

%Introduction to group here, including the overall objective of work in this area.
This working group concerned itself with identifying processes that are not adequately
covered by general software engineering. Verification and testing is
the first subtopic that we address.  
% there is a growing awareness about the lack of trust
% and reproducibility in many computational results. 
%\katznote{this paragraph repeats some of the previous one - they probably should be merged}
Computational science and engineering applications have many moving
parts that need to interoperate with one another. The accuracy and
reliability of results produced by scientific software depends not
only on the individual components behaving correctly, but also on the
validity of their interactions. 
% Therefore, a rigorous verification
% process and a robust testing regime are critical requirements for
% scientific software.  There have been instances where inadequate
% verification has resulted in publication of wrong results and later
% retraction of publications.
As scientific understanding grows,
the corresponding computational software models are refined, leading
to more complex codes. Increasing complexity makes them more prone to
defects, not only in individual code units, but also in interaction
among units. Therefore, a strong verification process combined with a
rigorous testing regime plays a critical role in the prevention of
generating incorrect scientific results. However, most science teams
struggle to find a good solution for themselves. Causes range from
lack of exposure to the practices, to distrust of adopting practices
because they do not meet the needs of the teams developing such
software.  The current focus of our effort is on testing because it is the first step
towards building a software processes that can lead to provenance and
reproducibility, the hallmarks of quality science.
